Spatial Analysis

### 2021-11-08 Scripts/spatial_analisis.Rmd
fixed grid

eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  filter(eez_name == "USA (East Coast)")

@@

grid_eez_sw_df %>% 
  mutate(approach = "state_waters") %>% 
  bind_rows(grid_eez_fp_df) %>%
  mutate(approach = ifelse(is.na(approach),"fishing_port",approach)) %>% 
  group_by(state,approach) %>% 
  summarise(n_grid = n()) %>% 
  spread(approach,n_grid) %>% 
  mutate(difference =state_waters-fishing_port,
         percentage_sw = round((difference/state_waters)*100)
         )

@@

### Previouslly created grids

grids <- my_path("D", "Spatial/grid_eez_fp", name = "grid_eez_fp_df.csv", read = T) %>%
  mutate(
    spatial = "fp"
  ) %>% 
  bind_rows(
    my_path("D", "Spatial/grid_eez_sw", name = "grid_eez_sw_df.csv", read = T)
  ) %>% 
  select(-spp) %>% 
  mutate(
    spatial = ifelse(is.na(spatial),"sw",spatial)
  ) %>% 
  distinct(.keep_all = T)

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp =="Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                 ifelse(year > 2001,"Today",NA)
                          )
                   )
    )
  )

@@

grid_eez_fp_sf <-
  grid_eez_sw_sf %>% 
  filter(!group %in% grid_eez_fp_sf$group) %>% 
  mutate(index = NA) %>% 
  bind_rows(grid_eez_fp_sf)

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(data = subset(grid_eez_fp_sf, index != "NA")) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    # scale_color_manual(values = state_pallet) +
    # scale_fill_manual(values = state_pallet) +
    scale_color_manual(values = c(state_pallet[3:4],state_pallet[6:13])) +
    scale_fill_manual(values = c(state_pallet[3:4],state_pallet[6:13])) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot() +
    geom_sf(data = subset(grid_eez_fp_sf, index = "NA"), color = "white") +
    geom_sf(data = subset(grid_eez_fp_sf, index != "NA"), aes(color = group),size = 0.1, alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    facet_wrap(~ group) +
    geom_point(data = rename(ports_df, group = landing_state), aes(x = lon, y = lat), color = "#E6A0C4") +
    theme(legend.position = "top") +
    # scale_color_manual(values = state_pallet,
    scale_color_manual(values = c(state_pallet[3:4],state_pallet[6:13]),
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

>>>>

eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  filter(eez_name == "USA (East Coast)") %>%

@@

my_path("R", "Partial",name = "grid_eez_sw_sf.csv", read = T) %>% 
  mutate(approach = "state_waters") %>% 
  bind_rows(my_path("R", "Partial/grid_fp",name = "grid_eez_fp_df.csv", read = T)) %>%
  mutate(approach = ifelse(is.na(approach),"fishing_port",approach)) %>% 
  group_by(state,approach) %>% 
  summarise(n_grid = n()) %>% 
  spread(approach,n_grid) %>% 
  mutate(difference =state_waters-fishing_port,
         percentage_sw = round((difference/state_waters)*100)
         )

@@

# US EEZ shapefile
eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")) %>% 
  st_transform(crs = 4326) %>% # 4326
  filter(Name == "USA (East Coast)") %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) #0.1 for paper

### Previously created grids

grids <- bind_rows(
  my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Paralichthys dentatus"),
  my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Stenotomus chrysops"),
  my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Centropristis striata")
) %>% 
  mutate(spatial = "sw") %>% 
  bind_rows(
    my_path("D", "Partial/grid_fp", name = "grid_eez_fp_df.csv", read = T)
  ) %>% 
  mutate(spatial = ifelse(is.na(spatial),"fp",spatial),
         spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                 ifelse(spp == "scup", "Stenotomus chrysops", 
                        "Centropristis striata"
                 )
    )
  ) %>% 
  distinct(.keep_all = T)

# unique(grids$spp)

# Ports data
ports_df <- my_path("R","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th") %>%
  distinct(port_name, .keep_all = T) # No need for duplication

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat","spp")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp == "Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                 ifelse(year > 2001,"Today",NA)
                          )
                   )
    )
  )

@@

grid_eez_fp_sf <- grid_eez_sw_sf %>% 
  filter(!group %in% grid_eez_fp_sf$group) %>% 
  mutate(index = NA) %>% 
  bind_rows(grid_eez_fp_sf) %>% 
  distinct(.keep_all = T)

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(data = subset(grid_eez_fp_sf, index != "NA")) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    # scale_color_manual(values = state_pallet) +
    # scale_fill_manual(values = state_pallet) +
    scale_color_manual(values = c(state_pallet[3:4],state_pallet[6:13])) +
    scale_fill_manual(values = c(state_pallet[3:4],state_pallet[6:13])) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot() +
    geom_sf(data = subset(grid_eez_fp_sf, index = "NA"), color = "white") +
    geom_sf(data = subset(grid_eez_fp_sf, index != "NA"), aes(color = group),size = 0.1, alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    facet_wrap(~ group) +
    # geom_point(data = rename(ports_df, group = state_postal), aes(x = lon, y = lat), color = "#E6A0C4") +
    theme(legend.position = "top") +
    # scale_color_manual(values = state_pallet,
    scale_color_manual(values = c(state_pallet[3:4],state_pallet[6:13]),
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)
